---
title: "03_analysis"
author: "Sglatt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r packages}
if (!require("lme4")) {
  install.packages("lme4")
  require("lme4")
}
if (!require("lmerTest")) {
  install.packages("lmerTest")
  require("lmerTest")
}
if (!require("sjPlot")) {
  install.packages("sjPlot")
  require("sjPlot")
}
if (!require("jmv")) {
  install.packages("jmv")
  require("jmv")
}
if (!require("dplyr")) {
  install.packages("dplyr")
  require("dplyr")
}
if (!require("psych")) {
  install.packages("psych")
  require("psych")
}
if (!require("tidyverse")) {
  install.packages("tidyverse")
  require("tidyverse")
}
if (!require("emmeans")) {
  install.packages("emmeans")
  require("emmeans")
}

# Cohen's d for familiar effect sizes from MLMs
lme.dscore <- function(mod, data, type) {
  if (type == "lme4") {
    mod1 <- lmerTest::lmer(mod, data = data)
    eff <- cbind(summary(mod1)$coefficients[, 4], summary(mod1)$coefficients[, 3])
  }

  if (type == "nlme") {
    eff <- cbind(summary(mod)$tTable[, 4], summary(mod)$fixDF$terms)
  }

  colnames(eff) <- c("t", "df")
  eff <- as.data.frame(eff)
  eff$d <- (2 * eff$t) / sqrt(eff$df)
  eff <- eff[-1, ]
  return(eff)
}
```

```{r folder to save output}
if (!dir.exists("03_Output")) {
  dir.create("03_Output")
}
```

# Data
Data for modality x time models; created in 01_data.Rmd
```{r mlm data}
# Continuous outcomes data
plf_data_filtered <- read.csv("01_Output/treatment_data_sg_2026-05-12.csv")

# Filter outcomes data to pre and post treatment (remove -9)
plf_data_treatment <- plf_data_filtered %>%
  filter(month %in% c(0, 3, 6, 12))

# Look at modality n's
table(plf_data_treatment[plf_data_treatment$month == 0, ]$modality_type)

# Exclude multi-modal participants
plf_data_treatment_sens <- plf_data_treatment %>%
  filter(modality_type != 3)

# Make a baseline dataset
plf_data_baseline <- plf_data_treatment %>%
  filter(month == 0)

plf_data_baseline_sens <- plf_data_baseline %>%
  filter(modality_type != 3)

# Confirm
table(plf_data_baseline$modality_type_new)
```

Data for initiation/engagement; created in 01_data.Rmd
```{r itt data}
plf_data_itt <- read.csv("01_Output/initiation_data_sg_2026-05-12.csv")
```

# Make MLM vars
```{r mlms vars}
plf_data_treatment_sens <- plf_data_treatment_sens %>%
  mutate(
    modality_type_new_f = as.factor(modality_type_new), # convert original to factor
    modality_f = factor(modality_type_new_f, levels = c(0, 1), labels = c("In-person", "Telehealth")),
    month_f = factor(month, levels = c(0, 3, 6, 12)),
    site_new_f = as.factor(site_new)
  )

# Make a baseline SI variable
ssi_bl <- plf_data_treatment %>%
  filter(month == 0) %>%
  dplyr::select(id, ssi)

# Add baseline SI to all data
plf_data_treatment_sens <- plf_data_treatment_sens %>%
  left_join(ssi_bl, by = "id", suffix = c("", "_bl"))
```

# Treatment initiation
```{r initiate raw}
# Rates of initiation among those assigned
table(plf_data_itt$modality_assign)
table(plf_data_itt$Attendance_clean)
21 + 31 + 24 + 25 # 101

table(plf_data_itt$modality_assign, plf_data_itt$Attendance_clean)
6 + 20 + 10 + 10 # 46
15 + 11 + 14 + 15 # 55
```

# Treatment engagement and completion
```{r engage raw}
# in-person session attendance
summary(plf_data_baseline_sens[plf_data_baseline_sens$modality_type_new == 0, ]$n_sessions_attended_clean)
sd(plf_data_baseline_sens[plf_data_baseline_sens$modality_type_new == 0, ]$n_sessions_attended_clean)

# telehealth session attendance
summary(plf_data_baseline_sens[plf_data_baseline_sens$modality_type_new == 1, ]$n_sessions_attended_clean)
sd(plf_data_baseline_sens[plf_data_baseline_sens$modality_type_new == 1, ]$n_sessions_attended)

# Comparison
wilcox.test(n_sessions_attended_clean ~ modality_type_new, data = plf_data_baseline_sens)

# Range and IQR
plf_data_baseline_sens %>%
  group_by(modality_type_new) %>%
  summarise(
    n = n(),
    median_sessions = median(n_sessions_attended_clean, na.rm = TRUE),
    IQR_sessions = IQR(n_sessions_attended_clean, na.rm = TRUE)
  )

# Raw count and % of those initiated, engaged, completed by modality
plf_data_baseline_sens %>%
  group_by(modality_type_new) %>%
  summarise(
    n = n(),
    initiated_n = sum(initiated_clean, na.rm = TRUE),
    initiated_pct = mean(initiated_clean, na.rm = TRUE) * 100,
    engaged_n = sum(six_sess_clean, na.rm = TRUE),
    engaged_pct = mean(six_sess_clean, na.rm = TRUE) * 100,
    completed_n = sum(completer_clean, na.rm = TRUE),
    completed_pct = mean(completer_clean, na.rm = TRUE) * 100
  )
```

# Multilevel models with orthogonal polynomial contrasts for time
# Test cubic effects in both groups separately
```{r test cubic time}
src_model_cubic_ip <- lmer(
  srcs17 ~ poly(as.numeric(month), 3) +
    poly(as.numeric(month), 3) * ssi_bl +
    site_new_f +
    (1 | id),
  data = subset(plf_data_treatment_sens, modality_f == "In-person")
)

src_model_cubic_th <- lmer(
  srcs17 ~ poly(as.numeric(month), 3) +
    poly(as.numeric(month), 3) * ssi_bl +
    site_new_f +
    (1 | id),
  data = subset(plf_data_treatment_sens, modality_f == "Telehealth")
)

tab_model(
  src_model_cubic_ip,
  src_model_cubic_th
)
```

# Models with linear and quadratic effects
```{r mlms orthogonal}
# plf_data_treatment_sens$month_factor <- factor(plf_data_treatment_sens$month, ordered = TRUE)
# contrasts(plf_data_treatment_sens$month_factor) <- contr.poly(length(unique(plf_data_treatment_sens$month_factor)))

# Include linear + quadratic only (not cubic)
src_model_xcubic <- lmer(
  srcs17 ~ poly(as.numeric(month), 2) * modality_f + # Linear and quadratic
    poly(as.numeric(month), 2) * ssi_bl + # interact polynomial terms with baseline SI
    site_new_f +
    (1 | id),
  data = plf_data_treatment_sens
)

summary(src_model_xcubic)
tab_model(src_model_xcubic) # first degree (1) = linear, second degree  (2) = quadradic

srce_model_xcubic <- lmer(
  srcse ~ poly(as.numeric(month), 2) * modality_f +
    poly(as.numeric(month), 2) * ssi_bl +
    site_new_f +
    (1 | id),
  data = plf_data_treatment_sens
)

summary(srce_model_xcubic)
tab_model(srce_model_xcubic)

srci_model_xcubic <- lmer(
  srcsi ~ poly(as.numeric(month), 2) * modality_f +
    poly(as.numeric(month), 2) * ssi_bl +
    site_new_f +
    (1 | id),
  data = plf_data_treatment_sens
)

summary(srci_model_xcubic)
tab_model(srci_model_xcubic)

tab_model(
  src_model_xcubic,
  srce_model_xcubic,
  srci_model_xcubic
) # In this table,first degree (1) = linear, second degree  (2) = quadradic

lme.dscore(src_model_xcubic, data = plf_data_treatment_sens, type = "lme4")
lme.dscore(srce_model_xcubic, data = plf_data_treatment_sens, type = "lme4")
lme.dscore(srci_model_xcubic, data = plf_data_treatment_sens, type = "lme4")

# Save the models
# saveRDS(src_model_xcubic, file = "03_Output/src_model_xcubic.rds")
# saveRDS(srce_model_xcubic, file = "03_Output/srce_model_xcubic.rds")
# saveRDS(srci_model_xcubic, file = "03_Output/srci_model_xcubic.rds")
```

# Test quadratic vs linear model fit (orthogonal polynomials)
```{r quad v linear}
# global coping
src_linear <- lmer(
  srcs17 ~ poly(as.numeric(month), 1) * modality_f +
    poly(as.numeric(month), 1) * ssi_bl + site_new_f + (1 | id),
  data = plf_data_treatment_sens
)
anova(src_linear, src_model_xcubic) 

# external coping
srce_linear <- lmer(
  srcse ~ poly(as.numeric(month), 1) * modality_f +
    poly(as.numeric(month), 1) * ssi_bl + site_new_f + (1 | id),
  data = plf_data_treatment_sens
)
anova(srce_linear, srce_model_xcubic)

# internal coping
srci_linear <- lmer(
  srcsi ~ poly(as.numeric(month), 1) * modality_f +
    poly(as.numeric(month), 1) * ssi_bl + site_new_f + (1 | id),
  data = plf_data_treatment_sens
)
anova(srci_linear, srci_model_xcubic)
```

# Center time & refit models
```{r center time}
plf_data_treatment_sens$month_num <- as.numeric(plf_data_treatment_sens$month)
plf_data_treatment_sens$month_centered <- plf_data_treatment_sens$month_num # baseline-centered
plf_data_treatment_sens$month2_centered <- plf_data_treatment_sens$month_centered^2

# Refit models with interpretable centered time
# Global SRC
src_model_final <- lmer(
  srcs17 ~ month_centered * modality_f +
    month2_centered * modality_f +
    month_centered * ssi_bl +
    month2_centered * ssi_bl +
    site_new_f + (1 | id),
  data = plf_data_treatment_sens
)

# External coping
srce_model_final <- lmer(
  srcse ~ month_centered * modality_f +
    month2_centered * modality_f +
    month_centered * ssi_bl +
    month2_centered * ssi_bl +
    site_new_f + (1 | id),
  data = plf_data_treatment_sens
)

# Internal coping
srci_model_final <- lmer(
  srcsi ~ month_centered * modality_f +
    month2_centered * modality_f +
    month_centered * ssi_bl +
    month2_centered * ssi_bl +
    site_new_f + (1 | id),
  data = plf_data_treatment_sens
)

summary(src_model_final)
summary(srce_model_final)
summary(srci_model_final)

tab_model(src_model_final)
tab_model(srce_model_final)
tab_model(srci_model_final)

tab_model(list(
  "Global SRC" = src_model_final,
  "External Coping" = srce_model_final,
  "Internal Coping" = srci_model_final
))

# Effect sizes
lme.dscore(src_model_final, data = plf_data_treatment_sens, type = "lme4")
lme.dscore(srce_model_final, data = plf_data_treatment_sens, type = "lme4")
lme.dscore(srci_model_final, data = plf_data_treatment_sens, type = "lme4")

# Save new models
saveRDS(src_model_final, "03_Output/src_model_final_centered.rds")
saveRDS(srce_model_final, "03_Output/srce_model_final_centered.rds")
saveRDS(srci_model_final, "03_Output/srci_model_final_centered.rds")
```

# Omnibus LRT
```{r omnibus LRT}
# Refit all models with ML
src_model_final_ml <- update(src_model_final, REML = FALSE)
srce_model_final_ml <- update(srce_model_final, REML = FALSE)
srci_model_final_ml <- update(srci_model_final, REML = FALSE)

# Global SRC LRT
src_null <- lmer(
  srcs17 ~
    month_centered +
    month2_centered +
    modality_f +
    month_centered * ssi_bl +
    month2_centered * ssi_bl +
    site_new_f +
    (1 | id),
  data = plf_data_treatment_sens,
  REML = FALSE
)

anova(src_null, src_model_final_ml)

# External SRC LRT
srce_null <- lmer(
  srcse ~
    month_centered +
    month2_centered +
    modality_f +
    month_centered * ssi_bl +
    month2_centered * ssi_bl +
    site_new_f +
    (1 | id),
  data = plf_data_treatment_sens,
  REML = FALSE
)


anova(srce_null, srce_model_final_ml)

# Internal SRC LRT
srci_null <- lmer(
  srcsi ~
    month_centered +
    month2_centered +
    modality_f +
    month_centered * ssi_bl +
    month2_centered * ssi_bl +
    site_new_f +
    (1 | id),
  data = plf_data_treatment_sens,
  REML = FALSE
)


anova(srci_null, srci_model_final_ml)
```

# Save analysis data
```{r save data}
# Save CSV
write.csv(plf_data_treatment_sens,
  file = paste0("03_Output/analysis_data_sg_", format(Sys.Date(), "%Y-%m-%d"), ".csv"),
  row.names = FALSE
)
```
